Spatial Reasoning and Statistical Graphics

### 2014-10-30 Paper/Draft1/TestingVisualAptitude.Rnw
Formatting changes to accommodate two-column layout

options(replace.assign=TRUE,width=70,scipen=3)

@@

score.summary <- melt(ans.summary, measure.vars=20:24, value.name = "Score", variable.name = "Test")
score.summary$Test <- factor(score.summary$Test, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure~Class.", "Paper Folding", "Visual Search"))
qrange$sqrt.k <- sqrt(qrange$k-1)
qrange$Test <- factor(qrange$testtype, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure~Class.", "Paper Folding", "Visual Search"))
ggplot() + 
  geom_bar(data=qrange, aes(width=.9, x=Test, y=-n/(k-1), fill="red"), alpha=.5, colour="red", stat="identity", inherit.aes=F) + 
  geom_bar(data=qrange, aes(width=.9, x=Test, y=n, colour="red"), stat="identity", fill=NA, inherit.aes=F) + 
  geom_boxplot(data=score.summary, aes(x=factor(Test), y=Score), inherit.aes=F, fill="grey") +
  scale_colour_discrete(name="", labels="Theoretical\nRange") +
  scale_fill_discrete(name="", labels="Scores\nindicating\nguessing") +
  theme_bw()  + 
#  theme(legend.position="bottom", legend.box="horizontal") + 
  ylab("Test Scores") + 
  xlab("") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  theme(plot.margin=unit(c(0,0,0,0),"mm"))
  #coord_flip()
@
%

@@

qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + 
  facet_wrap(~variable, scales="free_x", ncol=5) + xlab("") + 
  ylab("Scaled Lineup Score") + 
  geom_point(aes(x=value, y=lineup), shape=1, size=3) + 
  theme_bw() + 
  theme(plot.margins=unit(c(0,0,0,0), unit="mm"))
@
\begin{figure*}[hbpt]\centering
\includegraphics[width=.9\linewidth]{fig-VisReasoningCategorical-1}
\caption[Visual Aptitude Study Results]{Demographic characteristics of participants compared with lineup score. Categories are ordered by effect size; calculus completion,  hours spent playing video games per week, and sex are all associated with a difference in lineup score. }\label{fig:visualaptitudecat}
\end{figure*}

@@

<<correlations, echo=F, comment="|", warning=F, message=F, results='asis'>>=
print(xtable(cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")]), digits=c(1, rep(3, 5)), caption="Correlation matrix for the five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize)
@
<<pcatests,echo=F, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".45\\linewidth", fig.height=4.5, fig.width=4.5, warning=F, message=F>>=
pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
screeplot(pca)
@
<<pcarotation,echo=F, results='asis'>>=
print(xtable(summary(pca), caption="Importance of components in principal component analysis of the five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize)
print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize)
@

<<pca, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Plots of the principal components with observations. PC1 seems to represent a general ``visual intelligence'' factor (and thus is not represented here). Figure~classification is strongly related to lineups (PC2, PC3), and visual search is related to lineup performance as well (PC3). Paper folding does not seem to be strongly related to lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.\\label{fig:biplots}">>=
biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), adj=.75)
biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), adj=.75)
@

>>>>

options(replace.assign=TRUE,width=40,scipen=3)

@@

score.summary <- melt(ans.summary, measure.vars=20:24, value.name = "Score", variable.name = "Test")
score.summary$Test <- factor(score.summary$Test, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure~Class.", "Paper Folding", "Visual Search"))
qrange$sqrt.k <- sqrt(qrange$k-1)
qrange$Test <- factor(qrange$testtype, levels=c("lineup", "card_rot", "fig_class", "folding", "vis_search"), labels=c("Lineups", "Card Rotation", "Figure~Class.", "Paper Folding", "Visual Search"))
ggplot() + 
  geom_bar(data=qrange, aes(width=.9, x=Test, y=-n/(k-1), fill="red"), alpha=.5, colour="red", stat="identity", inherit.aes=F) + 
  geom_bar(data=qrange, aes(width=.9, x=Test, y=n, colour="red"), stat="identity", fill=NA, inherit.aes=F) + 
  geom_boxplot(data=score.summary, aes(x=factor(Test), y=Score), inherit.aes=F, fill="grey") +
  scale_colour_discrete(name="", labels="Theoretical\nRange") +
  scale_fill_discrete(name="", labels="Scores\nindicating\nguessing") +
  theme_bw()  + 
  ylab("Test Scores") + 
  xlab("") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1), legend.position=c(.75,.8)) +
  theme(plot.margin=unit(c(0,0,0,0),"mm"))
  #coord_flip()
@
%

@@

qplot(data=lineup.summary.categorical, x=value, y=lineup, geom="boxplot") + 
  facet_wrap(~variable, scales="free_x", ncol=5) + xlab("") + 
  ylab("Scaled Lineup Score") + 
  geom_point(aes(x=value, y=lineup), shape=1, size=3) + 
  theme_bw() + 
  theme(plot.margin=unit(c(0,0,0,0), unit="mm"))
@
\begin{figure*}[hbpt]\centering
\includegraphics[width=.9\linewidth]{fig-VisReasoningCategorical-1}
\caption[Visual Aptitude Study Results]{Demographic characteristics of participants compared with lineup score. Categories are ordered by effect size; calculus completion,  hours spent playing video games per week, and sex are all associated with a difference in lineup score. }\label{fig:visualaptitudecat}
\end{figure*}

@@

<<correlations, echo=F, comment="|", warning=F, message=F, results='asis', size="footnotesize">>=
print(xtable(cor(ans.summary[,c("lineup","card_rot", "fig_class", "folding", "vis_search")]), digits=c(1, rep(2, 5)), caption="Correlation matrix for the five tests."), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize", scalebox=.9)
@
<<pcatests,echo=F, fig.cap="Scree plot of principle component analysis of performance on the different test batteries.", out.width=".8\\linewidth", fig.height=3, fig.width=4.5, warning=F, message=F, size="footnotesize">>=
pca <- prcomp(ans.summary[,c("lineup", "card_rot", "fig_class", "folding", "vis_search")], scale=T, retx=T)
screeplot(pca)
@
<<pcarotation,echo=F, results='asis'>>=
print(xtable(summary(pca), caption="Importance of components in principal component analysis of the five tests.",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize, size="footnotesize")
print(xtable(pca$rotation, caption="PCA Rotation matrix for all five tests.",digits=2), sanitize.text.function=sanitize, sanitize.rownames.function=sanitize, sanitize.colnames.function=sanitize)
@

<<pca, fig.cap="", fig.width=4.5, fig.height=4.5, warning=F, message=F, out.width='.45\\linewidth', comment="|", fig.cap="Plots of the principal components with observations. PC1 seems to represent a general ``visual intelligence'' factor (and thus is not represented here). Figure~classification is strongly related to lineups (PC2, PC3), and visual search is related to lineup performance as well (PC3). Paper folding does not seem to be strongly related to lineup performance outside of the first principal component; card rotation is only positively associated with lineup performance in PC4.\\label{fig:biplots}", echo=F>>=
biplot(pca, choices=2:3, pc.biplot=T, cex=c(.5, 1), adj=.75)
biplot(pca, choices=4:5, pc.biplot=T, cex=c(.5, 1), adj=.75)
@
